---
title: "Data Preparation and EDA: Singapore Fertility Trends: A Visual Analytics Study"
author: "G1 T16 - Cheok En Qi and Teong Sze Ching"
date: "February 28, 2026" 
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  warning: false
  freeze: true
---

## 1.0 Getting Started

Install and launch the following R packages

```{r}
pacman::p_load(tidyverse, readxl, knitr, ggthemes, patchwork, janitor, lubridate, scales)
```

## 2.0 Data

The dataset is provided in Excel format and contains multiple time series. The worksheet includes both metadata rows and actual time-series observations. The raw file consists of 815 rows and 118 columns, indicating that further cleaning is required before analysis.

```{r}
raw0 <- read_xlsx("data/sg fertilty population.xlsx", col_names = FALSE)
cat("Raw shape:", nrow(raw0), "rows x", ncol(raw0), "cols\n")
```

## 3.0 Data Preparation

To prepare the dataset for analysis and visualisation, the raw worksheet was cleaned in several stages. Since the Excel file contains both metadata rows and actual time-series observations, the data first had to be restructured into an analysis-ready format. The preparation process involved extracting the original series names, identifying valid date rows, removing non-data rows, converting values into numeric format, and creating cleaned datasets for the shared EDA and later analytical tasks.

### 3.1 Extract series names (getting the variable names)

The first row of the worksheet contains the original series labels rather than actual data values/observations. This step extracts those labels, removes extra spaces, fills in any blank names with placeholders, and converts them into clean, unique variable names. A lookup table (series_key) is also created so that the cleaned names can still be traced back to the original labels in the dataset.

```{r}
series_labels <- raw0 %>%
  slice(1) %>%
  select(-1) %>%
  unlist(use.names = FALSE) %>%
  as.character() %>%
  str_squish()

blank_idx <- which(is.na(series_labels) | series_labels == "")
if (length(blank_idx) > 0) series_labels[blank_idx] <- paste0("series_", blank_idx)

series_names <- make.unique(janitor::make_clean_names(series_labels))
series_key <- tibble(series = series_names, label = series_labels)
```

Below shows a small sample of the original series labels and their cleaned variable names.

```{r,echo=FALSE}
sample_name_map <- tibble(original_label = series_labels, cleaned_name = series_names) %>%
  slice(1:10)

knitr::kable(sample_name_map)
```

### 3.2 Assign cleaned column names

After extracting the series labels, the first row is removed from the raw worksheet because it is no longer needed as data. The cleaned series names are then assigned as column names, while the first column is temporarily kept as time_raw so that it can be parsed into valid dates in the next step.

```{r}
raw1 <- raw0 %>%
  slice(-1) %>%
  setNames(c("time_raw", series_names))
```

### 3.3 Parse date rows and identify valid time-series observations (finding which rows are real dates and remove unnecessary rows)

The first column of the worksheet contains a mixture of date values and non-data text. To separate real observations from metadata rows, a custom date-parsing function is used. This function handles both standard date strings and Excel serial dates. Rows are then flagged as valid time-series rows only if the parsed date is not missing and the year falls within a reasonable range. This allows metadata and summary rows to be excluded from further analysis. 

```{r}
parse_ceic_date <- function(x) {
  if (inherits(x, "Date")) return(x)
  if (inherits(x, "POSIXt")) return(as.Date(x))

  x_chr <- str_squish(as.character(x))
  dt1 <- suppressWarnings(lubridate::ymd_hms(x_chr, tz = "UTC", quiet = TRUE))
  dt2 <- suppressWarnings(lubridate::ymd(x_chr, quiet = TRUE))
  d_text <- as.Date(dplyr::coalesce(dt1, dt2))

  num <- suppressWarnings(as.numeric(x_chr))
  use_serial <- !is.na(num) & num > 10000 & num < 60000
  d_serial <- as.Date(num, origin = "1899-12-30")

  dplyr::if_else(use_serial, d_serial, d_text)
}

raw1 <- raw1 %>% mutate(date = parse_ceic_date(time_raw))

is_ts_row <- !is.na(raw1$date) &
  lubridate::year(raw1$date) >= 1900 &
  lubridate::year(raw1$date) <= 2100

stopifnot(sum(is_ts_row) > 0)
```

### 3.4 Convert values into numeric format

After identifying the valid time-series rows, the dataset is reshaped into a tidy long format where each row represents one year, one series, and one value. The raw values are then converted into numeric format so that they can be used for plotting and analysis. Where multiple observations exist for the same series within the same year, only the latest non-missing observation is kept.

At this stage, two cleaned datasets are created: df_long_clean for flexible filtering and visualisation, and df_wide_clean for easier year-based comparison across variables.

The cleaned annual panel spans from 1950 to 2025 and contains 117 usable series, showing that the raw worksheet has been successfully transformed into an analysis-ready dataset.

```{r}
df_long_clean <- raw1 %>%
  filter(is_ts_row) %>%
  mutate(year = lubridate::year(date)) %>%
  pivot_longer(
    cols = all_of(series_names),
    names_to = "series",
    values_to = "value_raw"
  ) %>%
  mutate(value = readr::parse_number(as.character(value_raw))) %>%
  filter(!is.na(value)) %>%
  group_by(series, year) %>%
  slice_max(order_by = date, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  arrange(series, year)

df_wide_clean <- df_long_clean %>%
  select(year, series, value) %>%
  pivot_wider(names_from = series, values_from = value) %>%
  arrange(year)

cat("Annual panel year range:", min(df_wide_clean$year), "to", max(df_wide_clean$year), "\n")
cat("Non-missing series:", n_distinct(df_long_clean$series), "\n")



```

### 3.5 Create final dataset for EDA and Analytical tasks

This step prepares the final working datasets used in the rest of the report. This is necessary because the cleaned dataset still contains many series, while the report only uses selected subsets for specific charts and analyses.

After identifying the key fertility, marriage, and divorce series, they are combined into a core dataset, analysis_df, for the main analysis. This dataset keeps four columns: year, tfr, marriage, and divorce. Additional datasets are also created for the EDA charts, correlation analysis and regression analysis.

```{r}
pick_series <- function(key_tbl, primary_pattern, fallback_pattern = NULL) {
  out <- key_tbl %>%
    filter(str_detect(str_to_lower(label), primary_pattern)) %>%
    slice(1) %>%
    pull(series)
  if (length(out) == 0 && !is.null(fallback_pattern)) {
    out <- key_tbl %>%
      filter(str_detect(str_to_lower(label), fallback_pattern)) %>%
      slice(1) %>%
      pull(series)
  }
  out
}

tfr_series <- pick_series(series_key, "^fertility rate: per female$")
mar_series <- pick_series(series_key, "crude marriage rate", "marriage")
div_series <- pick_series(series_key, "crude divorce rate", "divorce")

stopifnot(length(tfr_series) == 1, tfr_series %in% df_long_clean$series)
stopifnot(length(mar_series) == 1, mar_series %in% df_long_clean$series)
stopifnot(length(div_series) == 1, div_series %in% df_long_clean$series)

analysis_df <- df_long_clean %>%
  filter(series %in% c(tfr_series, mar_series, div_series)) %>%
  select(year, series, value) %>%
  pivot_wider(names_from = series, values_from = value) %>%
  arrange(year) %>%
  transmute(
    year,
    tfr      = .data[[tfr_series]],
    marriage = .data[[mar_series]],
    divorce  = .data[[div_series]]
  )

tfr_df <- df_long_clean %>%
  filter(series == tfr_series) %>%
  arrange(year)

eda2_df <- df_long_clean %>%
  filter(series %in% c(tfr_series, mar_series, div_series)) %>%
  left_join(series_key, by = "series") %>%
  mutate(panel = case_when(
    series == tfr_series ~ "TFR (births per woman)",
    series == mar_series ~ "Marriage indicator",
    series == div_series ~ "Divorce indicator",
    TRUE ~ label
  ))

asfr_keys <- series_key %>%
  mutate(label_clean = str_to_lower(str_squish(label))) %>%
  filter(str_detect(label_clean, "fertility rate") &
           str_detect(label_clean, "1000") &
           str_detect(label_clean, "age|aged")) %>%
  mutate(
    age_band = str_extract(label, "\\d{2}\\s*[-–]\\s*\\d{2}"),
    age_band = str_replace_all(age_band, "\\s*", ""),
    age_band = str_replace_all(age_band, "–", "-")
  ) %>%
  filter(!is.na(age_band)) %>%
  distinct(series, .keep_all = TRUE)

age_levels <- c("15-19","20-24","25-29","30-34","35-39","40-44","45-49")

asfr_df <- df_long_clean %>%
  inner_join(asfr_keys %>% select(series, age_band), by = "series") %>%
  mutate(age_band = factor(age_band, levels = age_levels)) %>%
  filter(!is.na(age_band)) %>%
  arrange(year, age_band)

corr_df <- analysis_df %>%
  select(tfr, marriage, divorce) %>%
  cor(use = "complete.obs") %>%
  as.data.frame() %>%
  rownames_to_column("var1") %>%
  pivot_longer(-var1, names_to = "var2", values_to = "corr")

lag_cor <- function(x, y, max_lag = 10) {
  tibble(lag = 0:max_lag) %>%
    mutate(cor = map_dbl(lag, \(k) cor(x, dplyr::lag(y, k), use = "complete.obs")))
}

lag_results <- bind_rows(
  lag_cor(analysis_df$tfr, analysis_df$marriage, 10) %>% mutate(series = "Marriage leads TFR"),
  lag_cor(analysis_df$tfr, analysis_df$divorce, 10)  %>% mutate(series = "Divorce leads TFR")
)

lag_m <- 2
lag_d <- 1

analysis_lag <- analysis_df %>%
  mutate(
    marriage_lag = lag(marriage, lag_m),
    divorce_lag  = lag(divorce,  lag_d)
  ) %>%
  drop_na(marriage_lag, divorce_lag, tfr)
```

------------------------------------------------------------------------

## 4.0 EDA

Exploratory data analysis will be performed on the dataset to understand the main demographic trends and identify meaningful patterns before modelling.

The EDA will focus on three questions: when fertility began to decline, whether fertility, marriage, and divorce appear to move together over time, and whether fertility has shifted toward older age groups.

:::: callout-note
### 4.1 Trend of total fertility rate over time {.tabset}

::: panel-tabset
#### Visual

```{r, echo=FALSE}
ggplot(tfr_df, aes(x = year, y = value)) +
  geom_line() +
  geom_point(size = 1) +
  labs(
    title = "Singapore Total Fertility Rate (TFR) Over Time",
    x = "Year",
    y = "Births per woman"
  )
```

**Observation:**\
Singapore’s TFR shows a sustained long-run decline from the 1960s to 2025. The sharpest fall occurs from the 1960s through the late 1970s, followed by a slower downward drift with smaller fluctuations. From the mid-1970s onward, TFR remains well below the replacement level of 2.1, and the 2025 value of 0.87 is the lowest in the series, consistent with recent reports describing it as a historic low. A TFR of 0.87 means that, on average, a woman would be expected to have 0.87 children over her lifetime if the fertility rates observed in that year remained constant.

#### Code

``` r
ggplot(tfr_df, aes(x = year, y = value)) + geom_line() + geom_point(size = 1) + labs(title = "Singapore Total Fertility Rate (TFR) Over Time", x = "Year", y = "Births per woman")
```
:::
::::

------------------------------------------------------------------------

:::: callout-note
### 4.2 Local Fertility Peaks and Dragon-Year Comparison

::: panel-tabset
### Visual

```{r, echo=FALSE}
tfr_peaks <- tfr_df %>%
  select(year, value) %>%
  rename(tfr = value) %>%
  arrange(year) %>%
  mutate(
    local_bump = tfr - (lag(tfr) + lead(tfr)) / 2,
    is_peak = tfr > lag(tfr) & tfr > lead(tfr)
  ) %>%
  drop_na()

peak_years <- tfr_peaks %>%
  filter(is_peak) %>%
  arrange(desc(local_bump))

p1 <- ggplot(tfr_peaks, aes(x = year, y = tfr)) +
  geom_line(linewidth = 1) +
  geom_point(data = peak_years, size = 2) +
  labs(
    title = "Total Fertility Rate Over Time with Local Peaks Highlighted",
    x = "Year",
    y = "Births per woman"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"))

p2 <- ggplot(peak_years, aes(x = reorder(factor(year), local_bump), y = local_bump)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Local Bump in Peak Years",
    x = "Peak year",
    y = "Local bump"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"))

dragon_years <- c(1964, 1976, 1988, 2000, 2012, 2024)

peak_years_check <- peak_years %>%
  mutate(dragon_flag = if_else(year %in% dragon_years, "Dragon year", "Other year"))

p3 <- ggplot(peak_years_check, aes(x = reorder(factor(year), local_bump), y = local_bump, fill = dragon_flag)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Peak Years and Dragon-Year Overlap",
    x = "Peak year",
    y = "Local bump",
    fill = NULL
  ) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"))

p1 / p2
p3

```

### Code

``` r
tfr_peaks <- tfr_df %>%
  select(year, value) %>%
  rename(tfr = value) %>%
  arrange(year) %>%
  mutate(
    local_bump = tfr - (lag(tfr) + lead(tfr)) / 2,
    is_peak = tfr > lag(tfr) & tfr > lead(tfr)
  ) %>%
  drop_na()

peak_years <- tfr_peaks %>%
  filter(is_peak) %>%
  arrange(desc(local_bump))

p1 <- ggplot(tfr_peaks, aes(x = year, y = tfr)) +
  geom_line(linewidth = 1) +
  geom_point(data = peak_years, size = 2) +
  labs(
    title = "Total Fertility Rate Over Time with Local Peaks Highlighted",
    x = "Year",
    y = "Births per woman"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"))

p2 <- ggplot(peak_years, aes(x = reorder(factor(year), local_bump), y = local_bump)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Local Bump in Peak Years",
    x = "Peak year",
    y = "Local bump"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"))

dragon_years <- c(1964, 1976, 1988, 2000, 2012, 2024)

peak_years_check <- peak_years %>%
  mutate(dragon_flag = if_else(year %in% dragon_years, "Dragon year", "Other year"))

p3 <- ggplot(peak_years_check, aes(x = reorder(factor(year), local_bump), y = local_bump, fill = dragon_flag)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Peak Years and Dragon-Year Overlap",
    x = "Peak year",
    y = "Local bump",
    fill = NULL
  ) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"))

p1 / p2
p3
```
:::

**Observation:**\
Although TFR shows a clear long-run decline, several local peaks can still be observed in the series. To investigate these short-term increases, local peak years were first identified from the data, followed by a comparison with dragon years, which are commonly linked to higher birth numbers. The local bump measures how much a year’s TFR exceeds the average of the year before and after it. Hence, a value of 0.10 means the TFR is 0.10 births per woman above its immediate surrounding trend, while higher values indicate a stronger local spike.The analysis suggests that while there are some prominent peaks, such as 1976, 1988 and 2000, coincide with dragon years, but others do not. This suggests that dragon years may contribute to selected fertility spikes, but they do not explain all local fluctuations.
::::

------------------------------------------------------------------------

:::: callout-note
### 4.3 Fertility, Marriage and Divorce Trends {.tabset}

::: panel-tabset
#### Visual

```{r, echo=FALSE}
ggplot(eda2_df %>% filter(year >= 1980), aes(year, value)
) + geom_line() +  facet_wrap(~ panel, scales = "free_y", ncol = 1) + labs(title = "Fertility, Marriage and Divorce Trends", x = "Year", y = "Value")
```

**Observation:**\
The chart shows that from around 1980 onward, fertility continues to decline, marriage generally trends downward, and divorce generally trends upward before stabilizing. Together, these patterns suggest that fertility is changing alongside broader family-formation trends, although the chart alone does not establish causality.

Note: The chart is restricted to 1980 onward because the marriage and divorce indicators are only available from that period, making the overlapping years more suitable for comparison with TFR. The series do not end in exactly the same year because TFR is available up to 2025, while marriage and divorce are only available up to 2024 in the source dataset. This does not materially affect the analysis, since the multivariable comparison is based on the common years where all three indicators are available.

The marriage and divorce indicators refer to crude rates per 1,000 residents. For example, a marriage value of 6.5 means there were 6.5 marriages per 1,000 residents in that year, while a divorce value of 1.9 means there were 1.9 divorces per 1,000 residents.

#### Code

``` r
ggplot(eda2_df %>% filter(year >= 1980), aes(year, value)
) + geom_line() +  facet_wrap(~ panel, scales = "free_y", ncol = 1) + labs(title = "TFR vs Marriage/Divorce trend", x = "Year", y = "Value")
```
:::
::::

:::: callout-note
### 4.4 Age-specific fertility rate by age band over time {.tabset}

::: panel-tabset
#### Visual

```{r, echo=FALSE}
ggplot(asfr_df, aes(year, value, colour = age_band)) +  geom_line(linewidth = 0.9) + labs(title = "Age-Band Fertility Rates (Per 1,000 Females) Over Time", x = "Year", y = "Births per 1,000 females", colour = "Age band")
```

**Observation:**\
Across all age bands, fertility rates decline substantially over the long run, with the steepest drops occurring in younger age groups such as 15–19 and 20–24. Over time, the peak childbearing age shifts later: earlier decades are dominated by the 25–29 group, while from the 2000s onward the 30–34 group becomes the highest. The 35–39 group also remains relatively more persistent than younger groups in recent years. This pattern is consistent with birth postponement, where fertility shifts from women in their 20s toward women in their 30s, even as overall fertility continues to decline.

Note: Age-band series, such as 45–49, are not available for the full historical period in the source dataset. This does not materially affect the interpretation, as the chart is used to show the broad long-run shift in fertility across age groups.

The age-specific fertility rates are measured as births per 1,000 females within each age band. Thus, a value of 100 means there were 100 births per 1,000 females in that age group in that year.

#### Code

``` r
ggplot(asfr_df, aes(year, value, colour = age_band)) +  geom_line(linewidth = 0.9) + labs(title = "Age-Band Fertility Rates (Per 1,000 Females) Over Time", x = "Year", y = "Births per 1,000 females", colour = "Age band")
```
:::
::::

:::: callout-note
### 4.5 Fertility Postponement as Composition Shift (Age-Band Shares) {.tabset}

::: panel-tabset
#### Visual

```{r, echo=FALSE}

asfr_share_df <- asfr_df %>%
  group_by(year) %>%
  mutate(total_asfr = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(share = if_else(total_asfr > 0, value / total_asfr, NA_real_)) %>%
  filter(!is.na(share))

ggplot(asfr_share_df, aes(x = year, y = share, fill = age_band)) +
  geom_area() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    title = "Age-Band Fertility Composition Over Time",
    x = "Year",
    y = "Share of age-specific fertility",
    fill = "Age band"
  )
```

**Observation:**\
This chart shows the composition of age-specific fertility over time, where the shares sum to 100%. Over the decades, the share of fertility shifts away from younger age groups, especially 15–19 and 20–24, and gradually becomes more concentrated in older age groups, particularly 30–34 and 35–39. This provides clear visual evidence of birth postponement: even as overall fertility declines, the relative contribution of older age bands becomes more prominent.

Note: This chart complements the age-specific fertility rate chart in section 4.3 by showing how the relative contribution of each age band changes over time, thereby providing clearer evidence of fertility postponement.

#### Code

``` r
asfr_share_df <- asfr_df %>%
  group_by(year) %>%
  mutate(total_asfr = sum(value, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(share = if_else(total_asfr > 0, value / total_asfr, NA_real_)) %>%
  filter(!is.na(share))

ggplot(asfr_share_df, aes(x = year, y = share, fill = age_band)) +
  geom_area() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    title = "Age-Band Fertility Composition Over Time",
    x = "Year",
    y = "Share of age-specific fertility",
    fill = "Age band"
  )
```
:::
::::

------------------------------------------------------------------------

:::: callout-note
### 4.6 Scatterplots of TFR Against Marriage and Divorce

::: panel-tabset
### Visual

```{r, echo=FALSE}
p_mar <- ggplot(analysis_df, aes(x = marriage, y = tfr)) +
  geom_point(size = 2) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1) +
  labs(
    title = "TFR vs Marriage",
    x = "Marriage indicator",
    y = "TFR"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"))

p_div <- ggplot(analysis_df, aes(x = divorce, y = tfr)) +
  geom_point(size = 2) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1) +
  labs(
    title = "TFR vs Divorce",
    x = "Divorce indicator",
    y = "TFR"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"))

p_mar / p_div
```

### Code

``` r
p_mar <- ggplot(analysis_df, aes(x = marriage, y = tfr)) +
  geom_point(size = 2) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1) +
  labs(
    title = "TFR vs Marriage",
    x = "Marriage indicator",
    y = "TFR"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"))

p_div <- ggplot(analysis_df, aes(x = divorce, y = tfr)) +
  geom_point(size = 2) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1) +
  labs(
    title = "TFR vs Divorce",
    x = "Divorce indicator",
    y = "TFR"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold"))

p_mar / p_div
```
:::

**Observation:**\
The scatterplots suggest a positive association between TFR and the marriage indicator, and a negative association between TFR and the divorce indicator. Years with higher marriage values generally correspond to higher fertility, while years with higher divorce values tend to correspond to lower fertility. However, these patterns are descriptive and may partly reflect shared long-run trends over time rather than a direct relationship.
::::

------------------------------------------------------------------------

------------------------------------------------------------------------

:::: callout-note
### 4.7 Correlation Matrix of TFR, Marriage and Divorce

::: panel-tabset
### Visual

```{r, echo=FALSE}
corr_mat <- analysis_df %>%
  select(tfr, marriage, divorce) %>%
  cor(use = "complete.obs")

corr_df <- as.data.frame(as.table(corr_mat))

ggplot(corr_df, aes(x = Var1, y = Var2, fill = Freq)) +
  geom_tile(color = "white") +
  geom_text(aes(label = round(Freq, 2))) +
  labs(
    title = "Correlation Matrix ",
    x = NULL,
    y = NULL,
    fill = "Correlation"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

```

### Code

``` r
corr_mat <- analysis_df %>%
  select(tfr, marriage, divorce) %>%
  cor(use = "complete.obs")

corr_df <- as.data.frame(as.table(corr_mat))

ggplot(corr_df, aes(x = Var1, y = Var2, fill = Freq)) +
  geom_tile(color = "white") +
  geom_text(aes(label = round(Freq, 2))) +
  labs(
    title = "Correlation Matrix",
    x = NULL,
    y = NULL,
    fill = "Correlation"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```
:::

**Observation:**\
The correlation matrix shows that TFR is strongly positively correlated with marriage and strongly negatively correlated with divorce. Marriage and divorce are also strongly negatively correlated with each other. This suggests that higher fertility tends to be associated with higher marriage values and lower divorce values in the data.
::::

------------------------------------------------------------------------

```{r}
saveRDS(analysis_df, "data/analysis_df.rds")
saveRDS(tfr_df, "data/tfr_df.rds")
saveRDS(asfr_df, "data/asfr_df.rds")
saveRDS(series_key, "data/series_key.rds")
```
